---
title: "EDA Review with Penguins Data"
format: 
  revealjs:
    code-link: true
    chalkboard: true
    fig-align: center
    incremental: true 
    theme: dark
    smaller: true
    scrollable: true
editor: visual
execute: 
  echo: true
  warning: false
---

```{r packages}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(openintro)
library(janitor)
library(knitr)
library(kableExtra)
library(gridExtra)
library(readr)
library(gt)
library(palmerpenguins)
library(ggridges)
library(skimr)
```

## Palmer Penguins Data

::: nonincremental
-   We will explore the `penguins` data frame from the [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/) package. The data is the "size measurements for adult foraging penguins near Palmer Station, Antarctica." body_mass_g is the body mass of the penguin in grams
-   Below you see the first 10 rows of the dataframe
:::

```{r data}
head(penguins,10) |>
  gt()
```

## Palmer Penguins Data

-   How many penguins are in the dataframe?

-   How many variables were recorded about these penguins?

```{r data basics}
#| output-location: slide
glimpse(penguins)
```

## How many grams do the penguins tend to weigh?

-   What variable from the data frame will you need to use to answer this question?

-   What type of variable is it?

-   What is a visualization you could make to explore this question?

-   What summary statistics might you want to generate?

## How many grams do the penguins tend to weigh? 

### Visualization

-   Create a histogram to visualize the `body_mass_g` variable. Be sure to *add labels to each axis* and *a title*.

## How many grams do the penguins tend to weigh? 

### Visualization

```{r visualize weight}
penguins |>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram() +
  labs(
    title = "How many grams do penguins tend to weigh?",
    x = "Weight of Penguins in Grams",
    y = "Number of Penguins") +
  theme_minimal()
```

## How many grams do the penguins tend to weigh? 

### Statistics

::: nonincremental
-   Generate these statistics
:::

```{r weight statistics}
#| echo: false
penguins |>
  select(body_mass_g) |>
  skim()
```

## How many grams do the penguins tend to weigh? 

### Statistics

```{r weight statistics show}
penguins |>
  select(body_mass_g) |>
  skim()
```

## How many grams do the penguins tend to weigh? 

-   Answer the question by referencing the shape, center, and spread of the distribution.

## Is there a relationship between the species and weight?

-   What variables from the data frame will you need to use to answer this question?

-   What type is each variable?

-   What is a visualization you could make to explore this question?

-   What summary statistics might you want to generate?

## Is there a relationship between species and weight? 

### Visualize

-   Use the `facet_wrap` function to create a histogram for each species. (hint: copy and paste is your friend!)

-   The histograms should be positioned vertically. In other words you have a stack of 3 histograms.

-   Adjust the labs to match this new visualization

-   *Bonus:* Make each histogram a different color

## Is there a relationship between species and weight?

### Visualize

```{r weight species}
#| output-location: column
penguins |>
  ggplot(aes(x = body_mass_g, fill = species)) +
  geom_histogram() +
  labs(
    title = "How many grams do penguins tend to weigh?",
    subtitle = "Broken down by species",
    x = "Weight of Penguins in Grams",
    y = "Number of Penguins") +
  theme_minimal() +
  facet_wrap(~species, nrow = 3)
```

## Is there a relationship between species and weight?

### Statistics

-   Generate the necessary summary statistics to help you answer the question.

```{r stats weight species hidden}
#| echo: false
penguins |>
  group_by(species) |>
  select(body_mass_g) |>
  skim()
```

## Is there a relationship between species and weight?

```{r stats weight species}
penguins |>
  group_by(species) |>
  select(body_mass_g) |>
  skim()

```

## In which islands in the Palmer Archipelago were these penguins found?

-   What variable from the data frame will you need to use to answer this question?

-   What type of variable is it?

-   What is a visualization you could make to explore this question?

-   What summary statistics might you want to generate?

## In which islands in the Palmer Archipelago were these penguins found?

-   Create a table and a bar graph to explore this data.

-   Bonus: Have the table include the proportion of penguins located at each island.

## In which islands in the Palmer Archipelago were these penguins found? Table

```{r island table}
#| output-location: column
penguins |>
  count(island) |>
  mutate(proportion = (n/sum(n)))|>
  gt()
```

## In which islands in the Palmer Archipelago were these penguins found? Bar

```{r island bar}
#| output-location: column
penguins |>
  ggplot(aes(y = island, color = island)) +
  geom_bar() +
  labs(
    title = "islands in the Palmer Archipelago", 
    subtitle = "where penguins were found",
    x = "Number of Penguins"
  )
           
```
