---
title: "Name That Output and Write That Code"
format: 
  revealjs:
    code-link: true
    chalkboard: true
    fig-align: center
    incremental: true 
    theme: dark
    smaller: true
    scrollable: true
editor: visual
execute: 
  echo: true
  warning: false
---

```{r packages}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(openintro)
library(janitor)
library(knitr)
library(kableExtra)
library(gridExtra)
library(readr)
library(gt)
library(nflfastR)
library(nflreadr)
library(palmerpenguins)
library(ggridges)
library(skimr)
```

## Name that Output: Building Blocks

Tips:

-   Go line by line

-   Try to draw on paper or your iPad what will appear

```{r data frame}
#| code-line-numbers: "|1|2-4|5"
#| output-location: slide
rivers_data <- data.frame(
                    important_numbers = c(1915,333,6,12),
                    riv_colors = c("red", "white", "black", "gray")
                    )
rivers_data
```

## Name that Output: Visualize 1 Numerical Variable Description

-   It's winter so let's explore some data about penguins!

-   We will explore the `penguins` data frame from the [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/) package. The data is the "size measurements for adult foraging penguins near Palmer Station, Antarctica." body_mass_g is the body mass of the penguin in grams

-   Draw a sketch of what visualization the following code will produce.

```{r viz one var}
#| code-line-numbers: "|1|2|3|4|5"
#| output-location: slide

penguins|>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(binwidth = 250, fill = "orange", color = "black") +
  scale_x_continuous(limits = c(2500, 6500), breaks = seq(2500, 6500, 500))
```

## Write that Code: Upgrade the Visualization

```{r upgrade hist}
penguins|>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(binwidth = 250, fill = "orange", color = "black") +
  scale_x_continuous(limits = c(2500, 6500), breaks = seq(2500, 6500, 500))
```

## Write that Code: KEY Upgrade the Visualization

```{r upgrade hist key}
#| code-line-numbers: "|2-4|9-12|14|15-25"
#| output-location: slide
# Calculate the median value of total income
penguins_na <- penguins |>
                filter(!is.na(body_mass_g))
median_weight <- median(penguins_na$body_mass_g)

penguins|>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(binwidth = 250, fill = "orange", color = "black") +
    labs(x = "Body Mass of Penguins (grams)", 
       y = "Number of Penguins", 
       title = "How much do penguins tend to weigh?", 
       subtitle = "Data from a sample of 344 penguins near Palmer Station, Antarctica") +
  scale_x_continuous(limits = c(2500, 6500), breaks = seq(2500, 6500, 500)) +
  theme_minimal() +
  # Add median line with color and label
  geom_vline(aes(xintercept = median_weight), color = "navy", linetype = "dashed") +
  # Add text with median value next to the line
  annotate(
    "text",
    label = paste("Median:", round(median_weight, digits = 2)),
    x = median_weight + 100,
    y = 45,
    color = "navy",
    hjust = "left"
  )
```

## Name that Output: Visualize 1 Categorical Variable

-   Sketch what this code will produce.

-   What question(s) would cause you to create this visualization?

```{r bar}
#| code-line-numbers: "|1|2|3|4-7|8"
#| output-location: slide
penguins |>
  ggplot(aes(y= fct_rev(fct_infreq(species)), fill = species)) +
  geom_bar() +
  labs(x = "Number of Penguins", 
       y = "", 
       title = "Species of Penguins", 
       subtitle = "Data from a sample of 344 penguins near Palmer Station, Antarctica") +
  theme_minimal()
```

## Write that Code: Visualize 1 Categorical Variable with a Table

-   Write code that will create a table showing the number penguins in each species and the proportion of each species.

```{r viz for table}
#| echo: false
#| output-location: slide
penguins |>
  ggplot(aes(y= fct_rev(fct_infreq(species)), fill = species)) +
  geom_bar() +
  labs(x = "Number of Penguins", 
       y = "", 
       title = "Species of Penguins", 
       subtitle = "Data from a sample of 344 penguins near Palmer Station, Antarctica") +
  theme_minimal()
```

## Write that Code: KEY Visualize 1 Categorical Variable with a Table

```{r table}
#| code-line-numbers: "|1|2|3|4|5"
#| output-location: column
penguins |>
  count(species) |>
  mutate(proportion = n/sum(n)) |>
  arrange(desc(n)) |>
  gt()
```

## Write that Code: Generating Summary Statistics

-   What code would generate the summary statistics to support the following histogram.

```{r one var stats}
#| echo: false
penguins |>
  ggplot(aes(x= flipper_length_mm)) +
  geom_histogram()
```

## Write that Code: KEY Generating Summary Statistics

```{r sum stats one var}
penguins |>
  skim(flipper_length_mm)
```

## Write that Code: Generating Summary Statistics Across Groups

-   What did you learn from this visualization that you could not have learned from the previous one?

```{r two var stats}
#| echo: false
penguins |>
  ggplot(aes(x= flipper_length_mm, fill = species, color = species)) +
  geom_histogram(position = "identity", alpha = 0.5)
```

## Write that Code: KEY Generating Summary Statistics Across Groups

```{r sum stats groups}
penguins|>
  group_by(species) |>
  skim(flipper_length_mm)
```

## Challenge! Write that code!

```{r 3 var scatter}
#| echo: false
penguins |>
  ggplot(aes(x = flipper_length_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Flipper Length (mm)", 
       y = "Bill Length (mm)", 
       title = "What is the relationship between the length of a penguins\nflipper and bill length across three different penguin species?", 
       caption = "Data from a sample of 344 penguins near Palmer Station, Antarctica", 
       color = "Species") +
  theme_minimal()
```

## Challenge KEY! Write that code!

```{r 3 var scatter key}
#| code-line-numbers: "|2|3|4|5-10"
penguins |>
  ggplot(aes(x = flipper_length_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Flipper Length (mm)", 
       y = "Bill Length (mm)", 
       title = "What is the relationship between the length of a penguins\nflipper and bill length across three different penguin species?", 
       caption = "Data from a sample of 344 penguins near Palmer Station, Antarctica", 
       color = "Species") +
  theme_minimal()
```

## Sources:

<https://allisonhorst.github.io/palmerpenguins/>

<https://quarto.org/docs/presentations/revealjs/>

<https://github.com/OpenIntroStat/ims>

<https://sl.bing.net/jpUndVRcqSy>

<https://sl.bing.net/jLritSfGGia>
