---
title: "KEY_Tutorial 1: Introduction to R and RStudio"
author: "Mr. Adams"
editor: visual
format: 
  html:
    fig-width: 8
    fig-height: 4
---

# Welcome to R and RStudio!

Congratulations! You are taking the first step toward becoming a data scientist. R is a powerful statistical programming language and RStudio, the IDE (integrated development area), is the place where we will do our work with R.

This tutorial is the first in a series of tutorials that will help you begin to learn how to to explore data using R.

# By the end of this tutorial I should be able to...

## 1: Describe what a package is and know when to run a library.

If you were to build something, you would need to get some tools and organize those tools in various toolboxes. To build a house you'd need a variety of toolboxes because there are many different types of things you need to get done. For example, building the frame would require different set of tools than painting all of the rooms.

When working with R you can think of **packages as your toolboxes**. They are made up of various functions that make certain tasks easier. However, before using them, you need to go buy them and then pick them up off the shelf. Running the line `library(package_name)` is equivalent to grabbing your toolbox off the shelf before you start to work. If you don't grab the toolbox off of the shelf or, in other words, run the libraries, your code will not run.

*Long story short...run the libraries code first!*

`tidyverse` is a very robust package that we will use all of the time.

### 1.1: Tasks and Questions

1.  ***Run the following code by clicking the green play button.** (Note:You can run code by either clicking the green play button or highlighting the code you want to run and hitting command+return on your keyboard. The green play button will run the whole chunk while the keyboard option will mentioned above will only run the code you've selected.)*

```{r run libraries code, warning=FALSE}
library(tidyverse)
```

## 2: Use R as a calculator

At its most basic level, R can be used as a calculator. **Keep your phone in your pocket or backpack** and stay on task if you need to perform a calculation.

### 2.1: Tasks and Questions

1.  Below is an example of a calculation. In the gray space (aka code chunk) below, write your own calculation and then hit the play button.

2.  Write another calculation that requires you to add two numbers together and then divide that sum by 3.

```{r calculator}
#Example

3+2

#1 - Type your calculation below

20+14

#2 - Type your calculation below

(20+14)/3
```

## 3: Name functions

As you start to explore data, you will have many lines of code that create many statistics and visualizations. Staying organized is important. One way to do that is naming certain chunks of code. When you want to run that code again, you **run the name** of the code instead of the whole long chunk.

`name <- thing you want to name` *(Note: the \<- is made using the greater than and the minus sign)*

(Note: That little arrow is made by typing the greater than sign and minus sign on keyboard)

Naming tips:

-   use \_ instead of spaces. Example john_adams
-   keep it short yet something that will help you remember what it represents. Ex. `grades_hist` could be used to name a histogram showing the distributions of grades.

### 3.1: Tasks and Questions

1.  Copy one of the calculations you did in 2.1 Tasks and Questions and paste it into the gray section below.

2.  Click in front of the first number in your calculation and type in a name for your calculation and then add the arrow `<-`. name \<- calculation

3.  Type the name of your calculation on the line below what you did for question 2 and then run the code.

4.  Create and name 2 more calculations.

5.  Use the **names** to multiply all of your equations together.

```{r naming}
# 1,2,3

my_calc_two <- (20+14)/3
my_calc_two

#4
mult <- 1*2
mult

divide <- 8/2
divide

#5
my_calc_two*mult*divide

```

## 4: Create a vector

A vector is a sequence of either numbers or characters. Here are some examples.

**Here is an example of a numeric vector**

In this example, we create a vector of numeric values and then give it a name. `rivers_numeric_vector` is written on a line of its own so we can see a printout of the vector. Run the code to see what happens.

```{r numeric vectors }

rivers_numeric_vector <- c(1915,333,6,12)

rivers_numeric_vector
```

**In this example, we create a vector of characters and then give it a name.**

Remember, run the name of the vector by itself to see it.

```{r characters vectors}

rivers_character_vector <- c("red", "white", "black", "gray")

rivers_character_vector
```

### 4.1: Tasks and Questions

In the gray space below...

1.  Create and name a numeric vector that has 4 observations.

2.  Create and name a character vector that has 4 observations.

Bonus: Name the significance of each number in the vector created above titled, `rivers_numeric_vector`.

1915 - The year Rivers was founded.

333 - Rivers is located on 333 Winter St.

6 - Lowest grade at Rivers

12 - Highest grade

```{r create vectors}
# 1. Create and name our numeric vector in the line below. 

num_vector <- c(25,21,8,17)

# 2. Create and name your character vector in the line below.

char_vector <- c("a","b","c","d")

```

## 5: Create a data frame

As you've learned before, we will want our data frames organized as tidy datasets. When data is organized in such a way, each column is a variable and the rows are for each observational unit or case.

Most often, we will load in data to Rstudio or use data included within various packages. For this exercise, we will build a data frame.

Here is an example of a data frame. Use it to answer the tasks and questions section below.

```{r rivers data frame}
rivers_data <- data.frame(
                    important_numbers = c(1915,333,6,12),
                    riv_colors = c("red", "white", "black", "gray")
                    )
rivers_data
```

### 5.1: Tasks and Questions

1.  What is the name of this data frame? Type it EXACTLY how it appears in the code above.

    `rivers_data`

2.  How many variables are in this data frame and what are their names? Type the names EXACTLY how they appear in the code above.

    There are two variables in this data frame, `important_numbers` and `riv_colors`

3.  What types of variables are they?

    `important_numbers` is numeric

    `riv_colors` is categorical

4.  How many observations are in this data frame?

    4

### 5.2: Tasks and Questions

1.  In the gray space below, create and name a data frame that includes two variables and 4 observations.

```{r create a data frame}
# 1: Create and name your data frame in the lines below.

my_first_df <- data.frame(
                  num_vector = c(1,2,3,4),
                  char_vector = c("a","b","c","d")
)
my_first_df
```
